### 2017-06-05 abstract.tex
empty draft of paper

\documentclass[
    ,final            % use final for the camera ready runs
%%  ,draft            % use draft while you are working on the paper
%%  ,numberedheadings % uncomment this option for numbered sections
%%  ,                 % add further options here if necessary
    ,sort&compress
  ]
  {aipproc}

\layoutstyle{8x11double}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{amsmath}
\usepackage{units}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% FRONTMATTER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Paradigms in Physics 2.0}

\classification{} % <Replace this text with PACS numbers; choose from this list:
                  % \texttt{http://www.aip.org/pacs/index.html}>
\keywords{curriculum redesign} % <Enter Keywords here>

\author{David Roundy}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Liz Gire}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Ethan Minot}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Emily van Zee}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Tevian Dray}{
  address={Department of Mathematics, Oregon State
    University, Corvallis, Oregon 97331}}
\author{Corinne A. Manogue}{
  address={Department of Physics, Oregon State
    University, Corvallis, Oregon 97331}}

\newcommand\myderiv[3]{\ensuremath{\left(\dfrac{\partial #1}{\partial #2}\right)_{#3}}}
\newcommand\myderivinline[3]{\ensuremath{(\slfrac{\partial #1}{\partial #2})_{#3}}}
\newcommand\inexactd{d\hspace{-0.9ex}^-\!}

\newcommand\todo[1]{\textcolor{red}{(todo: #1)}}
\newcommand\fixme[1]{\textcolor{red}{(fixme: #1)}}
\newcommand\wording[2][red]{\textcolor{#1}{\underline{#2}}}
\newcommand\wordingnote[3][red]{\textcolor{#1}{\underline{#2}\footnote{#3}}}

\definecolor{small}{rgb}{0, 0, 1}

\newcommand\dbar{{\mkern6mu\mathchar'26\mkern-12mu d}}

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  This poster will describe both the process by which our
department achieved consensus on this curricular change, and the
resulting curriculum.  The Paradigms 2.0 committee begain with a
survey of students and faculty, followed by individual interviews with
the faculty teaching each existing course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include elimination or major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students to for their
upper-division courses.
\end{abstract}

\maketitle

%% \section{Introduction}\label{intro}

%% \begin{theacknowledgments}
%%   %% NSF DUE 0618877, 0837829, 1023120, and 1323800.
%%   The funding for this project was provided, in part, by the National
%%   Science Foundation under Grant Nos.  DUE 0618877, DUE 0837829,
%%   DUE 1023120, and DUE 1323800.
%% \end{theacknowledgments}

%% \bibliographystyle{aipproc}   % if natbib is available
%% \bibliography{paper}% Produces the bibliography via BibTeX.
% \input{paper.bbl}

\end{document}